---
title: "Homework 04"
output: html_document
---

```{r} 
library(tidyverse)
library(haven)
library(readxl)
library(tibble)
```

For this assignment, you'll need to open up, clean and save the following datasets, using the tools we've gone over in class. For each dataset, make sure that when you're done you have a nice, neatly labeled dataset that would be easy for you or another analyst to open and analyze. Save the result in an `RData` file using the names provided. You need to turn in an R script named `04-assignment_<lastname>.Rmd` that cleanly creates all of the datasets requested (lower case only for your last name, please).  

1. Panel data for OECD countries. Save as file name `oecd.RData` [http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat](http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat)


```{r}
oecd<-read.table("http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat",skip = 1)
 
names(oecd)<-c("country",
                "year",
                "lgaspcar",
                "lincomep",
                "lrpmg",
                "lcarpcap")

 
save(oecd,file="oecd.Rdata")
```



2. U.S. National Election Survey data from 2004. Save as file name `nes.RData`. [http://www.uta.edu/faculty/story/DataSets.htm](http://www.uta.edu/faculty/story/DataSets.htm)


```{r}
nes_spss<-read_spss("http://www.uta.edu/faculty/story/DataSets/NES2004.sav")

head(nes_spss)
```
Get rid of unwanted columns
```{r}
nes2<-nes_spss%>%select(-c(3))
```
Now save the file. 
```{r}
save(nes2,file="nes.Rdata")
```


3. General social survey, student version. Save as file name `gss.Rdata`. [http://www.uta.edu/faculty/story/DataSets.htm](http://www.uta.edu/faculty/story/DataSets.htm)

```{r}
if(file.exists("gss.xls")==FALSE){
download.file("http://www.uta.edu/faculty/story/DataSets/GeneralSocialSurvey1996ExcelDataStudentVersion.xls",destfile = "gss.xls")
gss<-read_excel("gss.xls",skip=4,col_names=FALSE)  
}else{
  	gss<-read_excel("gss.xls",skip=4,col_names=FALSE)
}
head(gss)

```

```{r}
#renaming columns
names(gss)<-c("id",	"hrs1",	"marital",	"childs",	"age",	"educ",	"sex",	"race",	"income",	"rincome",	"income91",	"partyid",	"polviews",	"libath",	"colrac",	"colmil",	"colhomo",	"cappun",	"relig",	"attend",	"reliten",	"bible",	"health",	"class",	"abdefect",	"abnomore",	"abhlth",	"abpoor",	"abrape",	"absingle", "abany",	"chldidel",	"homosex",	"xmovie",	"fear",	"fefam","eqincome",	"govtpow",	"aidold",	"aidunemp",	"aidcol",	"aidhouse",	"racseg",	"pornlaw")

```


4. Replication file for "STATISTICAL DISCRIMINATION OR PREJUDICE? A LARGE SAMPLE FIELD EXPERIMENT". Open up and save the `mainData.csv` file. Save it as `maindata.RData`. [Available here, look under the "Data and Analysis" tab.](http://thedata.harvard.edu/dvn/dv/restat/faces/study/StudyPage.xhtml?globalId=doi:10.7910/DVN/26410&studyListingIndex=1_f1a615c2cbee97fb79c9d575ae73)

```{r}
maindata<-read_csv(file="https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/26410/05NHUX")

write_csv(maindata,path="maindata.csv")

#Check it out
head(maindata)

##Need these for later
my.names<-names(maindata)

#Write this in a variety of formats to be used later
write_delim(maindata, path="maindata.txt",delim="\t")


gdata::write.fwf(data.frame(maindata),file="maindata.dat",sep="",colnames=FALSE)

save(maindata,file="maindata.Rdata")
```


5. The Lalonde dataset, covering work experiences in the Panel Study of Income Dynamics (psid). Save as `psid.RData` [http://users.nber.org/~rdehejia/data/psid_controls.txt](http://users.nber.org/~rdehejia/data/psid_controls.txt) You'll find a description of the data that you'll need [here](http://users.nber.org/~rdehejia/nswdata2.html).

```{r}
psid_stata1<-read_dta("http://www.nber.org/~rdehejia/data/psid_controls.dta")
 
head(psid_stata1)
 
psid<-psid_stata1%>%select(-c(data_id,1,7))
```

Getting rid of columns 
```{r}
psid2<-psid%>%select(-c(1,7))
```
 
Saving file 
```{r}
save(psid2,file="psid.Rdata")
```


